The function--sheaf dictionary for smooth commutative group schemes over finite fields

### 2014-04-03 dictionary.tex
Added a few todos for next time. Goodnight!

%%%%%%%%%%%%%%%%%% HEADING %%%%%%%%%%%%%%%%%%
\title[The function--sheaf dictionary]{The function--sheaf dictionary for smooth commutative group schemes over finite fields}
\date{\today}
\author{Clifton Cunningham}
\email{cunning@math.ucalgary.ca}
\address{University of Calgary}
\author{David Roe}
\email{roed.math@gmail.com}
\address{Pacific Institute for the Mathematical Sciences at the University of Calgary}
\subjclass[2010]{14F05 (primary), 14L15 (secondary)}
\keywords{sheaf--function dictionary, geometrization, quasicharacter sheaves, commutative group schemes}

@@

Motivated by ...., in this paper we ... The component group of the Neron model of algebraic torus over local field ...   The sheaf--function dictionary for etale group schemes is significantly different . . .\todo{Needs a new front end to really set the tone, and highten the contrast with the connected case, and point to applications.}

@@

Three examples will illustrate some of the issues that arise when extending the function--sheaf dictionary to smooth commutative group schemes.
First, consider the etale group scheme $\ZZ$ over $\Fq$.
While the $\ell$-adic character $\ZZ \to \EEx$ defined by
$1 \mapsto -1$, has finite order, the character defined by
$1 \mapsto 1+\ell$ does not, although its image is bounded;
on the other hand, the character $\ZZ \to \EEx$ defined by
$1 \mapsto \ell$ has unbounded image.
As we show in the paper, $\ell$-adic local systems on $G$ can only be used to geometrize characters of finite order; quasicharacter sheaves, however, provide the geometrization of all characters of $G(\Fq)$.
Second, consider the etale group scheme $\ZZ \times \ZZ$.
\todo{Build an example of an invisible quasicharacter sheaf.}
Finally, consider the etale group scheme $\ZZ \times \ZZ$ over $\Fq$ with Galois action determined by $\sigma(n,m) = (m,n)$ when the restriction of $\sigma$ to $\mathbb{F}_{q^2}$ is non-trivial.
Although there are infinitly-many geometric points on this group, it has exactly one $\Fq$-rational point.
In this case, the group carries only one isomorphism class of quasicharacter sheaves -- in particular, it has no invisible quasicharacter sheaves -- so the dictionary is perfect, albeit trivial, in this case.

@@

@@

\begin{proof}
Pick $\chi \in G(\Fq)^*$.
By Lemma~\ref{lem:dual-inv}, there is some $f\in G(\bFq)^*$ such that $f\vert_{G(\Fq)^*} = \chi$.
Set $\gqcs{L}_x = \EE$ for every $x\in G(\bFq)$.
Define $\mu_{x,y} : \gqcs{L}_{x+y} \to \gqcs{L}_x\otimes \gqcs{L}_y$ by $\mu_{x,y}(1) = 1 \otimes 1$.
Define $\phi_{x} : \gqcs{L}_{\Frob{}(x)} \to \gqcs{L}_x$ by $\phi_{x}(1) = f(x)$.
Since $f : G(\bFq) \to \EEx$ is a group homomorphism,
condition \eqref{nohom} is satisfied with $g =1$, the trivial cocycle in $Z^2(G(\bFq),\EEx)$.
It is now clear that we have defined a \emph{visible} quasicharacter sheaf $\qcs{L}$ with $\trFrob{\qcs{L}} = \chi$.
\end{proof}

@@

\begin{proof}

\end{proof}

>>>>

%%%%%%%%%%%%%%%%%% HEADING %%%%%%%%%%%%%%%%%%
\title[The function--sheaf dictionary]{The function--sheaf dictionary for smooth commutative group schemes over finite fields}
\date{\today}
\author{Clifton Cunningham}
\email{cunning@math.ucalgary.ca}
\address{University of Calgary}
\author{David Roe}
\email{roed.math@gmail.com}
\address{Pacific Institute for the Mathematical Sciences, University of Calgary}
\subjclass[2010]{14F05 (primary), 14L15 (secondary)}
\keywords{sheaf--function dictionary, geometrization, quasicharacter sheaves, commutative group schemes}

@@

Motivated by ...., in this paper we ... The component group of the Neron model of algebraic torus over local field ...   The sheaf--function dictionary for etale group schemes is significantly different . . .\todo{Needs a new front end to really set the tone, heighten the contrast to the connected case and point to applications.}

@@

Three examples will illustrate some of the issues that arise when extending the function--sheaf dictionary to smooth commutative group schemes.
First, consider the etale group scheme $\ZZ$ over $\Fq$.
While the $\ell$-adic character $\ZZ \to \EEx$ defined by
$1 \mapsto -1$, has finite order, the character defined by
$1 \mapsto 1+\ell$ does not, although its image is bounded;
on the other hand, the character $\ZZ \to \EEx$ defined by
$1 \mapsto \ell$ has unbounded image.
As we show in the paper, $\ell$-adic local systems on $G$ can only be used to geometrize characters of finite order; quasicharacter sheaves, however, provide the geometrization of all characters of $G(\Fq)$.
Second, consider the etale group scheme $\ZZ \times \ZZ$.
\todo{Build an example of an invisible quasicharacter sheaf here.}
Finally, consider the etale group scheme $\ZZ \times \ZZ$ over $\Fq$ with Galois action determined by $\sigma(n,m) = (m,n)$ when the restriction of $\sigma$ to $\mathbb{F}_{q^2}$ is non-trivial.
Although there are infinitly-many geometric points on this group, it has exactly one $\Fq$-rational point.
In this case, the group carries only one isomorphism class of quasicharacter sheaves -- in particular, it has no invisible quasicharacter sheaves -- so the dictionary is perfect, albeit trivial, in this case.

@@

\todo{Proof missing.}

@@

\begin{proof}
Pick $\chi \in G(\Fq)^*$.
By Lemma~\ref{lem:dual-inv}, there is some $f\in G(\bFq)^*$ such that $f\vert_{G(\Fq)^*} = \chi$.
Set $\gqcs{L}_x = \EE$ for every $x\in G(\bFq)$.
Define $\mu_{x,y} : \gqcs{L}_{x+y} \to \gqcs{L}_x\otimes \gqcs{L}_y$ by $\mu_{x,y}(1) = 1 \otimes 1$.
Define $\phi_{x} : \gqcs{L}_{\Frob{}(x)} \to \gqcs{L}_x$ by $\phi_{x}(1) = f(x)$.
Since $f : G(\bFq) \to \EEx$ is a group homomorphism,
condition \eqref{nohom} is satisfied with $g =1$, the trivial cocycle in $Z^2(G(\bFq),\EEx)$.
It is now clear that we have defined a \emph{visible} quasicharacter sheaf $\qcs{L}$ with $\trFrob{\qcs{L}} = \chi$.\todo{Can I improve this proof to show that the kernel of Trace of Frobenius is exactly $\Hh^0(\Weil{},\Hh^2(G(\bFq),\EEx))$? Perhaps . . . }
\end{proof}

@@

\begin{proof}
\todo{Clifton, you are here.}
\end{proof}
